---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

Loading the data and showing a summary of them and type of data

```{r, echo = TRUE}
activity <- read.csv("activity.csv",na.strings="NA")
summary(activity)
sapply(activity,class)
```

The field date is a factor which leves depends of each day of measurement

```{r, echo = TRUE}
levels(activity$date)
````

## Total number of steps taken per day

Using the library dplyr for grouping the data by day

```{r, echo = TRUE}
library(dplyr)
by_day <- group_by(activity, date)
steps_by_day <- summarise(by_day, steps = sum(steps, na.rm = TRUE))
steps_by_day
```

We plot the frequency of steps by day in a histogram way:

```{r, echo = TRUE}
hist(steps_by_day$steps)
```

Calculating mean and median of the total number of steps taken per day

```{r, echo = TRUE}
s1 <- summarise(steps_by_day, mean = mean(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))
s1
```

Therefore the mean of steps by day is `r s1$mean` and the median is `r s1$median`.

## Average daily activity pattern

To answer, we grup the data by interval and we take the average steps

```{r, echo = TRUE}
by_interval <- group_by(activity, interval)
avg_steps_by_interval <- summarise(by_interval, average = mean(steps, na.rm = TRUE))
avg_steps_by_interval
```

Ploting the average steps in each inetrval

```{r, echo = TRUE}
plot(avg_steps_by_interval$interval,avg_steps_by_interval$average, type = "l")
maxi <- max(avg_steps_by_interval$average)
interval <- avg_steps_by_interval[avg_steps_by_interval$average == maxi, 1]
```

The maximum average steps in an interval is `r maxi` and the correspondant interval is `r interval`

## Imputing missing values

```{r, echo = TRUE}
rows1 <- nrow(activity) 
rows2 <- nrow(activity[activity$steps == "NA",])
```

The number of rows in the activity data set is `r rows1` and the total number of records with missing values is `r rows2`. The next step is filling the missing values with the avereage of the interval

```{r, echo = TRUE}
activity2 <- mutate(activity, steps = replace(activity$steps, is.na(activity$steps), avg_steps_by_interval[avg_steps_by_interval$interval == c(0,5),]$average))
head(activity)
head(activity2)
by_day2 <- group_by(activity2, date)
steps_by_day2 <- summarise(by_day2, steps = sum(steps, na.rm = TRUE))
steps_by_day2
```

We plot the frequency of steps by day in a histogram for the new data set:

```{r, echo = TRUE}
hist(steps_by_day2$steps)
```

Calculating mean and median of the total number of steps taken per day

```{r, echo = TRUE}
s2 <- summarise(steps_by_day2, mean = mean(steps, na.rm = TRUE), median = median(steps, na.rm = TRUE))
s2
```

Therefore the mean of steps by day is `r s2$mean` and the median is `r s2$median` there is not a significant difference with the original data set

## Differences in activity patterns between weekdays and weekends

Creating the factor variable indicating weekdays and weekend days

```{r, echo = TRUE}
wday <- as.POSIXlt(activity$date)$wday
activity3 <- mutate(activity, wday = factor(sapply(wday, function (w) if (w  == 0 | w == 6) 0 else 1 ), levels=c (0,1), labels = c("weekend", "weekday")))
```

```{r, echo = TRUE}
by_interval3 <- group_by(activity3, wday ,interval)
avg_steps_by_interval3 <- summarise(by_interval3, average = mean(steps, na.rm = TRUE))
```

Ploting the average steps in each inetrval for ecah factor 

```{r, echo = TRUE}
par(mfrow = c(2, 1))
with(avg_steps_by_interval3, {
   plot(average ~ interval, data = avg_steps_by_interval3[wday == "weekend", ], type = "l", main = "weekend")
   plot(average ~ interval, data = avg_steps_by_interval3[wday == "weekday", ], type = "l", main = "weekday")
   })
```

As the grpah shows, saturday and sunday are more intenses in the number of steps.


